Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2017-02-15 draft.tex
.

\documentclass[10pt,a4paper]{article}
%\usepackage{layouts}
\usepackage[a4paper,margin=1.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{caption}

@@

In any case, modeling homeostasis includes an error estimation, which requires setting some predefined target that the feedback control will attempt to achieve. Of course this raises the question of how to choose these homeostatic targets. Experimental studies have found that neuronal populations exhibit a strong heterogeneity within their individual activity. Many statistical analyses found particular evidence for a log-normal like distribution of firing rates \cite{Buzsaki_Fir_Rates_2014,Wohrer_Fir_Rates_2012}. Following these results, one approach of modeling this feature would be to randomly distribute individual targets according to the aforementioned statistics. A problem of this procedure is the absence of any particular deeper physiological motivation or causation. In the light of this conceptual problem, one could, due to a lack of better knowledge, simply define a single global target of activity. It has been shown however that heterogeneity of neural response onto external input within a population is crucial for efficient stimulus encoding \cite{Marsat_2010}. Allowing for a broad distribution of activity thus has to be regarded an important compartment of models of cortical activity and structure.

A conceptual problem that one might consider is the fact that temporal fluctuations of activity are believed to be a major part of information transmission and processing within the brain \cite[p. 123-150]{Theor_Neur_Dayan}. Homeostasis as described above could potentially override information by eliminating these fluctuations of activity. However, it is argued that homeostatic mechanisms act on time scales of hours to days and thus do not interfere with changes in activity taking place on the order of seconds or below \cite{Turrigiano_2011}.

Returning to the previous question, though the slowness of homeostatic adaptation might allow for temporal fluctuations of activity, it does not overcome the conceptual dichotomy between single cell homeostasis and persistent variability in spontaneous activity. Marder and Goaillard speculated that instead of a mechanism controlling activity on a single-cell level, one might consider the possibility of a "global sensor of network performance to provide error signals for tuning of the network" \cite{Marder_2006}. This concept was used by Sweeney et al. as the basic idea for the introduction of a diffusive, intercellular homeostatic mechanism, which will be introduced in the next section.

>>>>

\documentclass[10pt,a4paper]{article}
%\usepackage{layouts}
%\usepackage[a4paper,margin=1.5in]{geometry}
\usepackage{geometry}
\geometry{
a4paper,
left=2.0in,
right=1.0in}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{caption}

@@

\subsubsection{The Role of Firing Rate Heterogeneity -  a Dichotomy between Heterogeneity and Control?}
Modeling homeostasis necessarily includes a form of error estimation, which requires setting some predefined target that the feedback control will attempt to achieve. Of course this raises the question of how to choose these homeostatic targets. Experimental studies have found that neuronal populations exhibit a strong heterogeneity within their individual activity. Many statistical analyses found particular evidence for a log-normal like distribution of firing rates \cite{Buzsaki_Fir_Rates_2014,Wohrer_Fir_Rates_2012}. Following these results, one approach of modeling this feature would be to randomly distribute individual targets according to the aforementioned statistics. A problem of this procedure is the absence of any particular deeper physiological motivation or causation. In the light of this conceptual problem, one could, due to a lack of better knowledge, simply define a single global target of activity.

However, numerous experimental and theoretical studies have suggested that the presence of both slow- and fast-firing cells is not to be regarded as a ignorable, mere side effect of brain dynamics. It rather seems to be that skewed and heavy-tailed firing rate statistics are necessary with respect to preserving stable synaptic structures of long-term memory against relatively fast synaptic rewiring due to changes in external stimuli. For example, experiments by Dragoi et al. have found that during remapping of hippocampal place cells, neurons with low firing rates are subject to stronger changes within their field of spatial representation than those with high firing rates. In this context Buzsáki and Mizuseki argue that a small subset of highly active neurons may serve to form relatively strongly connected and stable sub-networks that can quickly react to momentary changes in the environment, while fine tuned adaptation over longer time periods involves the slower firing majority of cells \cite{Buzsaki_Fir_Rates_2014}. A realistic network model of neural activity and plasticity should therefore allow for a broad distribution of firing rates.

Interestingly, measurements in the barrel cortex and hippocampus have found that firing rates during exploration or task performance are significantly correlated with those found for spontaneous, baseline activity \cite{OConnor_2010,Mizuseki_2013}. This suggests that the same set of neurons tend to be relatively active or inactive under different stimuli and changing environments, in line with the idea of modeling homeostasis by means of individual targets of activity, which together form the desired statistical distribution. Still, though apparently being a good phenomenological approach, it is unclear what gives rise to these target differences.

Another problem that one might consider is the fact that temporal fluctuations of activity are believed to be a major part of information transmission and processing within the brain \cite[p. 123-150]{Theor_Neur_Dayan}. Homeostasis as described above could potentially override information by eliminating these fluctuations of activity. However, it is argued that homeostatic mechanisms act on time scales of hours to days and thus do not interfere with changes in activity taking place on the order of seconds or below \cite{Turrigiano_2011}. Returning to the previous question, though the slowness of homeostatic adaptation might allow for temporal fluctuations of activity, it does not overcome the conceptual dichotomy between single cell homeostasis and persistent variability in spontaneous activity. Marder and Goaillard speculated that instead of a mechanism controlling activity on a single-cell level, one might consider the possibility of a "global sensor of network performance to provide error signals for tuning of the network" \cite{Marder_2006}. This concept was used by Sweeney et al. as the basic idea for the introduction of a diffusive, intercellular homeostatic mechanism, which will be introduced in the next section.
